About

### 2023-08-07 tutorial/02-feature-extraction.qmd
tutorial updated

For our illustration, we rely on the work of researchers in the ESM literature, where different research teams have proposed a number meaningful measures to capture time series features that are relevant to psychological time series [@kuppens2017;@krone2018;dejonckheere2019;@wendt2020]. The main article discusses the individual time series features, their conceptual interpretation, as well as the technical calculation in more detail. Here we focus on the code to extract the features for the included variables.

The selection of time series features should always be driven by the research questions and field-specific conventions, but for our illustration we chose six time series features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity. We provide the `featureExtractor()` function to extract a number of measurements for each time series aspect. The time series feature operationalizations we chose for our illustration here are marked in bold.

>>>>

For our illustration, we rely on the work of researchers in the ESM literature, where different research teams have proposed a number meaningful measures to capture time series features that are relevant to psychological time series [@kuppens2017;@krone2018;@dejonckheere2019;@wendt2020]. The main article discusses the individual time series features, their conceptual interpretation, as well as the technical calculation in more detail. Here we focus on the code to extract the features for the included variables.

The selection of time series features should always be driven by the research questions and field-specific conventions, but for our illustration we chose six time series features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity. We provide the `featureExtractor()` function to extract a number of measurements for each time series aspect. The function is also available as part of the R package that bundles a number of functions we used for the illustration ([Download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz)). The time series feature operationalizations we chose for our illustration here are marked in bold.

### 2023-08-07 tutorial/05-evaluation.qmd
tutorial updated

```{r}
#| label: tbl-cluster-performance
#| fig-cap: Cluster Performance for the k = 2–10

@@

df_kmeans_performance %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = "c",
      digits = 2,
      row.names = FALSE,
      col.names = kmeans_metrics_names) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%")
```

@@

```{r}
#| label: fig-clusterFeatVarComb
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 18
#| fig-cap: Note that n and discrimination are out of cluster variables.

@@

We discuss the individual differences and their theoretical relevance as part of the main manuscript. However, in terms of methodological insight, we find variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail. In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2). We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics [@bringmann2018c;eronen2021a].

@@

```{r}
#| label: fig-cluster-timeseries
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 18

@@

```{r}
#| label: fig-mad-cluster-comparison
#| message: false
#| warning: false

@@

feature_plot_list$mad

feature_ttest_apa(results_df, "mean", "EvDayDiscr.post", "html")
```

>>>>

```{r}
#| label: tbl-cluster-performance
#| tbl-cap: Cluster Performance for k = 2–10

@@

df_kmeans_performance %>%
  kbl(.,
      escape = FALSE,
      booktabs = TRUE,
      align = "c",
      digits = 2,
      row.names = FALSE,
      col.names = kmeans_metrics_names) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%")
```

@@

```{r}
#| label: fig-clusterFeatVarComb
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 18
#| fig-cap: Cluster comparison by feature and variable. Note that n and discrimination are out of cluster variables.

@@

We discuss the individual differences and their theoretical relevance as part of the main manuscript. However, in terms of methodological insight, we find variables and features that distinguish the clusters better than others and a combination of variables and features lets us explore meaningful group differences in more detail. In our case, we see that the central tendency, variability, and linear trend are best at distinguishing a group with mainly positive experiences (cluster 1) from a group with a more negative experience (cluster 2). We also see that our clusters line up with the past literature on the importance of focusing on simpler and more meaningful statistics [@bringmann2018c;@eronen2021a].

@@

```{r}
#| label: fig-cluster-timeseries
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 18
#| fig-cap: Average cluster trajectories (using LM and GAM)

@@

```{r}
#| label: fig-mad-cluster-comparison
#| message: false
#| warning: false
#| fig-cap: Cluster comparison MAD across variables

@@

feature_plot_list$mad
```

### 2023-08-07 tutorial/about.qmd
tutorial updated

We provide all data necessary to replicate the full results of the main manuscript as well as this illustration. We specifically provide a minial reproducible dataset that includes all variables necessary to perform the analysis, while still protecting the privacy and confidentiality of the participants. All data is available as part of [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].

@@

The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com) [@KreienkampMasked-clust-github]. A minimal version of the code as well as an archive file are available in [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf]. For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package. This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz) [@KreienkampBlinded2023-tsFeatureExtracR].

>>>>

We provide all data necessary to replicate the full results of the main manuscript as well as this illustration. We specifically provide a minial reproducible dataset that includes all variables necessary to perform the analysis, while still protecting the privacy and confidentiality of the participants. All data is available as part of [our Open Science repository](https://osf.io/j8dzv/?view_only=c765efa51e6545e588b29999611eb789) [@KreienkampMasked-clust-osf].

@@

The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com) [@KreienkampMasked-clust-github]. A minimal version of the code as well as an archive file are available in [our Open Science repository](https://osf.io/j8dzv/?view_only=c765efa51e6545e588b29999611eb789) [@KreienkampMasked-clust-osf]. For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package. This package is available as its own GitHub repository and can be downloaded with the `devtools` package `devtools::install_github("JannisCodes/tsFeatureExtracR")`, you can also [download the R package here](https://github.com/JannisCodes/tsFeatureExtracR/archive/refs/tags/v.0.1.0.tar.gz) [@KreienkampBlinded2023-tsFeatureExtracR].
